Fast response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks.

### 2015-08-20 manuscript/manuscript_wtc3.Rmd
edits to manuscript
deleted some test files on repo

small edits to plots

# Summary {.unnumbered}

* overview
* method
* big result
* implication

@@

\
First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @duursma2006vertical; @warren2007internal]. Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis]. Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves has major implications for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.

@@

\
2.	Under comparable leaf VPD, g~s~ should be proportional to A across both sun and shade leaves and g~m~ should scale positively with photosynthetic capacity.

@@

\
Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 L pots and kept inside each chamber. This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments. Seedlings were planted into each chamber after mean seedling height reached 100 cm.  The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils with low organic matter and low water holding capacity. A root exclusion barrier extended from chamber walls to the hard layer (ca. 1 m) and roots were allowed to grow freely below the barrier. After 2 months, floors were installed 45 cm from the soil surface.  This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored (for methods see Barton et al. [-@barton2010whole]). Leaf gas exchange measurements were then initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured.  At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^. For ET treatments trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.

@@

\
Once the chamber environment for leaves was stable, for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet. These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL. Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired WTC leaf, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference and samples gases for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle. Photosynthesis, g~s~, transpiration, vapor pressure deficit and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The ternary effect corrections (t) are described by:

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment. Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous water use efficiency (~i~WUE) was calculated as leaf photosynthesis divided by transpiration. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].

@@

## Photosynthetic capacity and leaf photosynthesis rates 
The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b). As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~, AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~. When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).

@@

## Stomatal conductance and leaf water use efficiency
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004). Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a). With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.

\
Leaf ~i~WUE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001). Following an increase in available light, ~i~WUE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001). ~i~WUE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2). For all leaf types and light treatments there was a strong response of ~i~WUE to VPD, and individual data points broadly corresponded to response curves from the optimal ~i~WUE model (Figure `r waterusefig`a). The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04. Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

@@

## Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.

@@

## Variation in CO~2~ drawdown
More open stomata in shade leaves, possible as a result of a non-limiting supply of leaf K, lead to a significant increase in C~i~ compared to sun leaves (Figure `r ciccfig`a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

# Discussion
Here we show that dynamic physiological responses in shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy. Tree canopies may maintain the ability to upregulate A beyond the constraints of their local light environment, even at the cost of valuable resources. Empirical data accounting for the total conductance of CO~2~ reveal that shifts in g~s~ and g~m~ in shade leaves can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought. Thus, failure to integrate the effects of g~m~ into process based tree growth models may underestimate the ability of canopies to alter their photosynthetic capacity when environmental conditions change. Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.

@@

\
Unexpected higher rates of g~s~ led to decreased instantaneous WUE in shade leaves throughout the experiment. Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across leaf lifespans. From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.  As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE while the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.

## Leaf physiology of sun and shade leaves
\
The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.  In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory. However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE. This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].  As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.

\
It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks. Rapid increases in leaf temperature with sunflecks has the potential to inhibit C gain [@leakey2003high]. It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal]. Here, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity. Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown. Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution]. Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy (see De Pury & Farquhar, [-@de1997simple]).

\
Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A. This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained. In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

## Conclusions
\
Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment. If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution. If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves which experience high amounts of sunflecks may be large enough to tolerate excess losses from transpiration. However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].

\
Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain.

The observed physiological behaviour of leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete. Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.  Additional empirical data across multiple tree species are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite lower mean PPFD [@pearcy1990sunflecks].

# List of Tables

**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean(standard error) from AC~i~ curves measured in each chamber at saturating light. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and warming treatments.

\
**Table 2**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* trees between full sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD~leaf~ are kPa. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and both light and warming treatments.

# List of Figures

**Figure 1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.

\
**Figure 2**. Local light environment for sun and shade leaves for each gas exchange campaign. Means with standard errors represent 1 m integrated PPFD (n=5), measured with a ceptometer, at the canopy height of each selected leaf.

@@

\
**Figure 8**. The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light with standard error

@@

# Tables {.unnumbered}

**Table 1**.

@@

# Figures {.unnumbered}
```{r airvars, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/air_vars.R")
```

@@

```{r acicurves, fig.width=6, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```

@@

```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```

@@

```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
source("master_scripts/wateruse.R")
```

>>>>

# Summary
* Light gradients within tree canopies play a major role in the distribution of resources that define photosynthetic capacity of individual leaves. The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis, stomatal conductance and mesophyll conductance of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers. The physiological behaviour of individual leaves was then correlated with the distribution of nitrogen and water within the canopy to evaluate how trees are optimized for carbon gain.
* More open stomata and rapid increases in mesophyll conductance allowed shade leaves to "lie in wait" for sunflecks and lead to leaf carbon gain exceeding sun leaves.
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. Evidence that mesophyll conductance not only varies within a canopy but can be up-regulated over short time intervals must now be considered in process based tree growth models.

# Key words
photosynthesis, stomatal conductance, mesophyll conductance,  shade, leaf optimal behavior

@@

\
First, g~s~ should be efficiently distributed within a canopy to utilize supplies of light, N and water to maximize A [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Second, variation in photosynthetic capacity between sun and shade leaves has also been hypothesized to be due to differences in g~m~ [@piel2002effect; @duursma2006vertical; @warren2007internal]. Mesophyll conductance can impose limitations on A as large as those associated with g~s~ [@warren2008stand; @ubierna2011estimation] and these limitations can reduce the efficiency of N use in A [@niinemets2007photosynthesis]. Interpreting the coupled responses of both g~s~ and g~m~ to A across sun and shade leaves is critical for predicting canopy C gain, but empirical measurements across tree canopies are still lacking.

@@

\
2.	Stomatal conductance should be proportional to A across sun and shade leaves under similar leaf VPD and g~m~ should scale positively with photosynthetic capacity.

@@

\
Before seedlings were planted into each chamber they were maintained under well watered conditions in 35 L pots and kept inside each chamber. This allowed for seedlings to gain sufficient biomass before planting while also allowing them to acclimate to chamber temperature treatments. Seedlings were planted into each chamber after mean seedling height reached 100 cm.  The top soils at this site, used in both pots and chambers, are an alluvial formation of low-fertility sandy loam soils with low organic matter and low water holding capacity. A root exclusion barrier extended from chamber walls to the hard layer (ca. 1 m) and roots were allowed to grow freely below the barrier. After 2 months, floors were installed 45 cm from the soil surface. This enabled chamber fluxes of CO~2~ and H~2~O from the whole tree canopy to be monitored (for methods see Barton et al. [-@barton2010whole]). Leaf gas exchange measurements were initiated in October 2013 when trees had both ample height growth and canopy development for realistic canopy light gradients to be measured. At this point, trees under AT treatment had a mean diameter of 28.2&pm;1.1 mm, height of 348&pm;15.1 cm and an estimated leaf area of 3.9&pm;0.1 m^2^. For ET treatments trees had a mean diameter of 34.1&pm;2.1 mm, height of 418.3&pm;23.1 cm and an estimated leaf area of 6.2&pm;0.2 m^2^.

@@

\
Once the chamber environment for leaves was stable, for each leaf measurement, the sample and reference gas lines were diverted to the TDL via T-junctions inserted into the reference gas tube and match valve outlet. These gases were dried by passing through napion gas dryers in the respective gas lines, and then ^12^CO~2~ and ^13^CO~2~ concentrations were measured for each gas by the TDL. Reference, sample and 2 calibration gases were run on alternating 80 s loops (20 s each), one for each paired WTC leaf, for a total of 12 min. This allowed for 4-5 measurements per leaf and data were averaged over the last 10 s of reference and samples gases for calculations. The two calibration gases were drawn from compressed air tanks (330 and 740 ppm CO~2~) in order to correct for gain drift of the TDLAS on each measurement cycle. Photosynthesis, g~s~, transpiration, VPD and intercellular [CO~2~] were auto logged every 15 s for each gas exchange system over the same 12 min interval.

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ is the fractionation from boundary layer diffusion (2.9&permil;) and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The ternary effect corrections (t) are described by:

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C for one sun and shade leaf for each chamber prior to the initiation of the drought treatment. Each AC~i~ curve began at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l l^-1^ at 25 &deg;C at saturating light (1800 $\mu$mols m^-1^ s^-1^). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@Duursma2014] in R [@RDevelopmentCoreTeam2011].

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous water use efficiency (WUE~i~) was calculated as leaf photosynthesis divided by transpiration. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].

@@

## Photosynthetic capacity and leaf photosynthesis rates 
The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Within individual chambers Vc~max~ was positively related to leaf N~a~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b). As g~m~ (results below) is included in the estimates of J~max~ and Vc~max~ with conventional AC~i~ curves, AC~c~ curves were simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~. When accounting for the effect of g~m~, chloroplastic Vc~max~ and J~max~ were still higher in sun leaves (Figure `r accfig`).

@@

## Stomatal conductance and leaf water use efficiency
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004). Increases in g~s~ were postively correlated to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a). With increased light availability, increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.

\
Leaf WUE~i~ was significantly greater in sun leaves than in shade leaves at low light (P = 0.001). Following an increase in available light, WUE~i~ of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001). WUE~i~ in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2). For all leaf types and light treatments there was a strong response of WUE~i~ to VPD, and individual data points broadly corresponded to response curves from the optimal WUE model (Figure `r waterusefig`a). The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04. Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

@@

## Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.

@@

## Variation in intercellular and chloroplastic CO~2~ concentrations
More open stomata in shade leaves under low and high light lead to significant increases in C~i~ compared to sun leaves (Figure `r ciccfig`a). The drawdown of CO~2~ from intercellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature]. This drawdown was similar between sun and shade leaves at their local light environment and increased marginally in shade leaves at high light (Figure `r ciccfig`c). This was the result of the proportional relationship between g~m~ and A across all leaves. The CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

# Discussion
Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment. More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to leaf C gain exceeding sun leaves. Although these results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies still impedes our ability to predict whole canopy C gain. Failure to integrate this dynamic physiology into process based tree growth models may underestimate the ability of canopies to alter leaf C gain when environmental conditions change. Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.

@@

\
Unexpected higher rates of g~s~ led to decreased WUE~i~ in shade leaves throughout the experiment. Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across the leaf lifespan. From a canopy perspective this pattern in WUE initially appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~.  As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE when the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.

## Physiological behaviour of sun and shade leaves
The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.  In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory. However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE. This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].  As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.

\
It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks. Rapid increases in leaf temperature with sunflecks has the potential to inhibit C gain [@leakey2003high]. It has been suggested that transpiration-induced cooling is effective in avoiding sunfleck induced heat damage, however leaves need to keep stomata open in the shade for this to occur [@schymanski2013stomatal]. Here, shade leaves were able to increase A equivalent or even outperforming sun leaves under identical light intensity. Whether more open stomata is a strategy to reduce stomatal response time or to decrease thermal load on shade leaves in trees is still unknown. Unfortunately, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution]. Thus, our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy [see @de1997simple].

\
Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A. This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained. In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of g~m~ [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

\
Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case the physiological adjustment of CO~2~ drawdown into shade leaves, is possibly directed towards occasional maximums of light availability over adaptation to a low light environment. If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution. If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves which experience high amounts of sunflecks may be large enough to tolerate excess losses from transpiration. However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].

## Conclusions
Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy. Although the distribution of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. These findings suggest that current theories of leaf optimal behaviour may be incomplete and have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized. Additional empirical data across multiple tree species are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

# List of Tables
**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, leaf N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and warming treatments.

\
**Table 2**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* trees between full sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across all gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and both light and warming treatments.

# List of Figures
**Figure 1**. Daily maximum and minimum temperature (a) and total daily PPFD (b) for each chamber across the experiment duration.

\
**Figure 2**. Local light environment for sun and shade leaves for each gas exchange campaign. Means &pm; 1 standard error represent 1 m integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.

@@

\
**Figure 8**. The mean &pm; 1 standard error of intercellular CO~2~ concentration (a), CO~2~ concentration in the chloroplasts (b) and CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light.

@@

# Tables {.unnumbered}
**Table 1**.

@@

# Figures {.unnumbered}
```{r airvars, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/air_vars.R")
```

@@

```{r acicurves, fig.width=8, fig.height=8, warning=FALSE,echo=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/aci_vcmaxnitro.R")
```

@@

```{r Agmgs, fig.width=5, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```

@@

```{r wateruse, fig.width=5, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=400}
source("master_scripts/wateruse.R")
```
